---
title: "Programming Base R"
format: html
editor: visual
---

# Task 1

## Q1 and Q2

We will define our pre and post treatment blood pressure **vectors**, ensuring we label the patients accordingly.

```{r}
pre_bp <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)
names(pre_bp) <- paste("Subject", 1:20, sep = "_")

post_bp <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
names(post_bp) <- paste("Subject", 1:20, sep = "_")

```

## Q3 and Q4

We will save these differences (pre-post) as a **vector**. We are curious what the **mean** difference is.

```{r}
diff_bp <- pre_bp - post_bp
mean(diff_bp)
```

## Q5

Which subjects saw a **decrease** in blood pressure?

```{r}
names(which(diff_bp>0))
```

## Q6 and Q7

What is the **mean decrease** among these subjects?

```{r}
mean(diff_bp[diff_bp>0])
```

# Task 2

## Q1

Lets create an **data frame** using the above **vectors**.

```{r}
bp_df <- data.frame(patient=names(pre_bp),pre_bp, post_bp, diff_bp, row.names = NULL)

```

## Q2

Which of the patients in our **data frame** saw an **increase** in blood pressure?

```{r}
bp_df[bp_df$diff_bp<0,]
```

## Q3 and Q4

Lets add a column to check whose post blood pressure is **less** than 120, and display our data frame.

```{r}
bp_df$post_less120 <- bp_df$post_bp<120
knitr::kable(bp_df)
```

# Task 3

## Q1

We are doing the exact same thing we did with the treatment group except with the placebo group to create a **data frame**.

```{r}
pre_bp2 <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
names(pre_bp2) <- paste("Subject", 1:10, sep = "_")
post_bp2 <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
names(post_bp) <- paste("Subject", 1:10, sep = "_")
diff_bp2 <- pre_bp2-post_bp2
bp_df2 <- data.frame(patient=names(pre_bp2),pre_bp2, post_bp2, diff_bp2, row.names = NULL)
bp_df2$post_less120 <- bp_df2$post_bp<120
knitr::kable(bp_df2)

```

## Q2

Now we will stack our two **data frames**, one for treatment one for placebo in a **list**.

```{r}
bp_list <- list(treatment=bp_df, placebo=bp_df2)
```

## Q3

We need to obtain the first element in our **list** three different ways.

-   Method 1:

```{r}
bp_list[[1]]

```

-   Method 2:

```{r}
bp_list[1]
```

-   Method 3:

```{r}
bp_list$treatment
```


## Q4

Finally we need to access the pre blood pressure placebo column of data in our above **list**.

```{r}
bp_list$placebo$pre_bp2

```

